---
title: "Package Upload: MATH 4753"
author: "Payton Schulte"
date: "`r Sys.Date()`"
output:   
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
vignette: >
  %\VignetteIndexEntry{Package Upload}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> This package contains functions and data from the course MATH-4753 Fall 2025 semester. 

# 1. ntickets() function

This function calculates how many airline tickets should be sold to keep the probability of overbooking to a certain risk level. The inputs for the function are:

* N - the number of seats on the plane
* gamma - the probability of overbooking
* p - the probability a passenger arrives

The outputs for the function are:

* nd - the optimal number of tickets to sell using discrete method
* nc - the optimal number of tickets to sell using continuous method
* N - the number of seats on the plane
* p - the probability a passenger arrives
* gamma - the probability of overbooking

```{r}
MATH4753F25paytonsch::ntickets(N = 100, gamma = 0.5, p = 0.95)
```

# 2. chebyshev() function

This function calculates the minimum proportion of data that must fall within k standard deviations of the mean according to Chebyshev's Theorem. The input for the function is:

* k - a numerical value greater than 1 representing number of standard deviations from the mean

The output for the function is:

* The minimum proportion of data within k standard deviations of the mean

```{r}
MATH4753F25paytonsch::chebyshev(2)
```

# 3. myboot2() function

This function performs a bootstrap simulation, plots the result, and calculates a confidence interval. The inputs for the function are:

* iter - the number of bootstrap resamples
* x - data vector
* fun - type of statistic, ie. mean, median
* alpha - significance level for confidence interval
* cx - plot size
* ... - any other arguments

The outputs for the function are:

* ci - bootstrap confidence interval
* fun - the statistic used
* x - the original data vector

```{r}
x <- rnorm(30)
print(MATH4753F25paytonsch::myboot2(iter=10000, x = x, fun="mean", alpha=0.05, cx=1.5))
```

# 4. myncurve() function

This function plots a normal density curve and shades the area under the curve to value a. The inputs for the function are:

* mu - the mean of the normal distribution
* sigma - the standard deviation
* a - the upper bound of the shaded region

The outputs for the function are:

* mu - the mean of the normal distribution
* sigma - the standard deviation
* a - the upper bound of the shaded region
* probability - the probability of X being less than a

```{r}
MATH4753F25paytonsch::myncurve(mu = 10, sigma = 2, a = 8)
```
